Beneficial arthropod habitat useage changes within-season: implications for pest control services in croplands

### 2020-11-04 manuscript/manuscript.Rmd
Worked on 2nd round of revisions

---
title: "Beneficial arthropod habitat useage changes within-season: implications for pest control services in croplands"
#Potential titles: 
# Non-crop features influence the abundance and timing of beneficial arthropod abundance
# 1) Cropland ecosystem services in space and time: Beneficial arthropod habitat use shifts within-season.
# 2) Beneficial arthropod habitat use shifts within-season: implications for pest control services in croplands
# 
author:
  - name: Samuel Robinson
    email: samuel.robinson@ucalgary.ca
    affiliation: University of Calgary
    footnote: 1
  - name: Diane Edwards
    email: diane.edwards.ca@gmail.com
    affiliation: Ambrose University
  - name: Jess Vickruck
    email: jessica.vickruck@canada.ca
    affiliation: University of Calgary
  - name: Lincoln Best
    email: lincoln.best@oregonstate.edu
    affiliation: Oregon State University
  - name: Paul Galpern
    email: paul.galpern@ucalgary.ca
    affiliation: University of Calgary
address:
  - code: University of Calgary
    address: 2500 University Drive NW, Calgary, AB, Canada
  - code: Ambrose University
    address: 150 Ambrose Circle SW, Calgary, AB
  - code: Oregon State University
    address: 1500 SW Jefferson Way, Corvallis, OR, USA
footnote:
  - code: 1
    text: "Corresponding Author"
date: "Spring 2020"
abstract: "Ground-dwelling arthropods are important generalist predators in agro-ecosystems, and can use non-crop features as overwintering habitats. However, it is unclear which types of landscape features constitute useful non-crop habitat, and at what spatial scale organisms gather resources at. Additionally, the same landscape feature may act as source of arthropods spillover at different times of the year, but this is rarely considered. We modeled the activity-density of four highly-abundant species of Canadian prairie arthropods, and used functional regression to simultaneously consider both their habitat preferences and the timing of sources and sinks. _Pterostichus melanarius_ (Coleoptera: Carabidae) and _Pardosa moesta_ (Arachnida: Lycosidae) appear to be attracted to canola (_Brassica napus_ L.) during the early summer, then disperse to grasslands, wetlands, and grassy road margins during the end of the summer. This suggests that grasslands and wetlands act as reservoir habitats for predatory arthropods, and that farmers should consider preserving existing habitat around croplands in order to maintain background levels of pest-control."

@@

# Introduction
Arthropods are one of the most influential groups of animals in agricultural systems.
Pest arthropods can cause large reductions in crop yield [@oerke2005], which can result in negative downstream economic effects and food insecurity.
On the other hand, spillover of beneficial arthropods from adjacent non-crop patches can deliver increased crop yields by providing ecosystem services such as extra pollination or pest suppression [@losey2006; @klein2007; @garibaldi2013].
Seasonal "pulses" of food resources from crop fields act as a bottom-up driver of their associated arthropod communities, meaning that agricultural land can act as a source of nutrients and individuals for adjacent non-crop land [@diekotter2010; @galpern2017].
These important fluxes of mobile arthropods, both to crops and away from them, are key for managing both crop yield and biodiversity in agroecosystems.
Finding "win-win" practices is of particular importance to land managers and conservationists; that is, practices that result in better conservation outcomes while maintaining or improving crop yields.
Finally, agricultural land occupies 15 million km$^2$ (~36%) of land globally [@ramankutty2018], suggesting that there remains great potential to achieve conservation outcomes by adjusting agricultural practices.

Uncultivated semi-natural land (SNL) can act as important habitat for beneficial arthropods [@duelli2003], which may spill over into adjacent cropland.
Therefore, landscape management is a strategy to increase beneficial arthropod abundance in crops [@albrecht2010; @landis2000; @martin2020; but see @tscharntke2016; @karp2018].
For example, wild bee abundance in agricultural landscapes is largely driven by flower and nest availability [@roulston2011], and the effect of SNL on bee abundance and diversity is generally positive.
Predatory arthropods such as beetles or spiders can be limited by water, food, and egg-laying sites [@lovei1996; @purtauf2005; @gardiner2010], but the effect of SNL on their abundance and diversity is less consistent [@macfadyen2013; @shackelford2013] as some taxa may use SNL as a travel corridor or overwintering site, but are not completely dependent on it for reproduction.
Some taxa are not dependent on SNL at all, and thrive in cultivated landscapes, while others exclusively inhabit SNL and do not venture beyond the edge of cultivated fields.
For example, carabid beetles are thought to reproduce in field margins, while the adults feed in fields [@desender1988].
This depends on the life history of the organisms in question, such as the seasonal timing of emergence, feeding, and reproduction, as well as their traits such as dispersal capability.
Non-adjacent cropland can also influence arthropod abundance if the organisms are long-distance dispersers [@oberg2008].
<!-- Small patches of SNL can have a large effect if overall landscape lacks SNL (landscape contrast [@boetzl2020; @vanSchalkwyk2020]) -->
In this way, the spillover of beneficial arthropods into crops depends not only on the amount of adjacent SNL, but the specific habitat type, the time of the year, and the spatial scale at which an organism gathers resources.

@@

In this study, we used pitfall traps to determine how landscape composition affects the seasonal activity-density of arthropods across a large Canadian prairie agroecosystem.
We considered the following hypotheses:
1. Untilled semi-natural land provides egg-laying and feeding areas for predatory arthropods, meaning that it should act as a source of arthropods during the early part of the season, and a sink during the later part of the season. 
2. Crops act as sources of food (pest insects) for predatory arthropods during the growing season. This should result in agricultural land becoming a sink for predators in the early part of the season as they migrate into the crop, and a source at the end of the season as they migrate out of the crop.
3. Crops may act as a temporary feeding site for predators. Therefore, crops may act as a local sink for arthropods, but will also be negatively associated at larger (landscape-level) scales.
Using a large pitfall trapping dataset, we related seasonal changes in arthropod abundance to landscape composition at multiple spatial scales.

@@

## Data collection
We used a set of `r sum(NtrapsPer$n)` pitfall traps installed in road margins (minimum of 5 m away from the road edge, `r with(NtrapsPer, sum(n[trapLoc=='ditch']))` traps) and in-field locations (`r with(NtrapsPer, sum(n[trapLoc!='ditch']))`) across southern Alberta, Canada in 2017 (Figure \@ref(fig:siteMap)).
The sites spanned a west-to-east gradient of four natural subregions, including foothills parkland, foothills fescue, mixedgrass, and dry mixedgrass [@ABRegions2006].
Traps were placed starting on `r format(as.Date(as.character(min(PteMelMod$tempTrap$startjulian)),format='%j'),format='%B %d')`, and collections ended on `r format(as.Date(as.character(max(PteMelMod$tempTrap$endjulian)),format='%j'),format='%B %d')`, with collection occurring continuously and traps being emptied every `r round(mean(PteMelMod$tempTrap$trapdays),1)` days on average (SD: `r round(sd(PteMelMod$tempTrap$trapdays),1)`). 
This resulted in `r nrow(PteMelMod$tempTrap)` unique collection events, taking place across a total of `r round(sum(PteMelMod$tempTrap$trapdays),1)` trapping days.
In-field traps were placed in canola crops (`r with(NtrapsPer, sum(n[trapLoc=='canola']))`), wetlands (`r with(NtrapsPer, sum(n[trapLoc=='wetland']))`), grassy field edges (`r with(NtrapsPer, sum(n[trapLoc=='pivot']))`), and remnant prairie grasslands (`r with(NtrapsPer, sum(n[trapLoc=='native']))`).
Traps in canola were installed at 25, 75, and 200 m along a transect heading away from the nearest non-crop feature (wetland, grassy field edges, or remnant prairie), while the trap at 0 m was installed in the non-crop feature itself.
We used 582 mL Solo® cups buried up to the rim and partially filled with ethylene glycol, with 2 cm wire mesh mounted over the rim to prevent vertebrates from falling into the traps.
Specimens were identified to species using appropriate taxonomic literature [@dondale1990; @edgar1990; @vogel2004; @yigit2007].

We used counts of organisms in our pitfall traps as proxies of ecosystem service provision.
However, increased counts of organisms in pitfall traps can represent higher activity levels (same number of organisms but more mobile) or higher abundances in the vicinity of the trap.
This makes it impossible to disentangle arthropod activity from density using single traps [@lang2000; @zhao2013], meaning that counts in the traps represent the "activity-density" of a given organism.
Despite this disadvantage, activity-density is generally positively related to ecosystem services, at least in carabids [@trichard2014; @boetzl2018; @gonzalez2020], making it acceptable for our study.

```{r siteMap, echo=FALSE, fig.cap='Location of pitfall traps during 2017, and classified cover f. Inset map shows study location within Alberta. Roadside traps were installed in road margins, while in-field traps were installed directly in canola fields or in adjacent wetland or grassy field edges.', out.width='100%'}  
  include_graphics('../figures/maps/PF2017Map.png',dpi=NA)
```

We used four highly-abundant species of predatory arthropods found in the pitfall traps, as common species are often more important for ecosystem service provision than rare species [@winfree2015], but their habitat preferences are still poorly understood.
_Pterostichus melanarius_ is an introduced predatory ground beetle (Coleoptera: Carabidae) that is a wide-ranging generalist known to prey upon crop pests [@carcamo1994; @larsen2003; @busch_Msc].
It is commonly found in untilled grasslands [@purtauf2005] and may use grassy field edges as larval habitat [@desender1988].
_Pardosa_ are wolf spiders (Araneae: Lycosidae) that use an active wandering predation strategy [@young1990], and are common across the Canadian prairies.
Both _Pardosa distincta_ and _Pardosa moesta_ are found in a wide variety of habitats, but are common in ungrazed pastures [@dondale1990; @carcamo2014].
_P. distincta_ is also commonly found in disturbed environments [@collins1996; @wade2010], while _P. moesta_ appear to prefer less frequent disturbances [@graham2003; @kowal2011; but see @dondale1990].
Other _Pardosa_ are known to use road margins and wooded areas as habitat [@buddle2000; @drapela2011].
Finally, _Phallangium opilio_ is an introduced harvestman (Opiliones: Phalangiidae) that is found in drier disturbed areas [@bragg2009].
Juveniles hatch from eggs during the spring, becoming adults during the summer, and eggs are laid during the fall [adults do not overwinter; @bragg2009].
They are commonly found in human-altered landscapes [@muster2014; @vandepoel2015Msc] and are nocturnal generalist hunters and scavengers [@halaj2000; @allard2005; @allard2005b], mainly eating small soft-bodied invertebrates [@acosta2007].

@@

Functional regression [@ramsay2004; @yen2014] was used to incorporate landscape information at different distances, allowing assessment of both local and regional landscape composition [@galpern2020].
Scalar-on-function regression is a special type of linear regression model ($y=X\beta + e$), where the columns of the model matrix $X$ contain some continuous predictor of the scalar $y$, and the values of the coefficients ($\beta$) are modeled as a smooth function $f(x)$ of the predictors.
<!-- As an example of scalar-on-function regression, biomass in a plot (scalar) could be modeled as a smoothed function of monthly average temperatures at that plot (function), meaning that the effect of monthly temperatures would also be a function of time. -->
In our case, the proportion of landscape cover ($X$) within each annulus surrounding the trap is a function of distance away from the trap, meaning that the coefficients are a smoothed function of distance ($\beta_i = f(distance_i)$), and represent the additive effect of a given type of landscape cover at distances away from the trap (i.e. $X\beta$).
This allows for the possibility that the size of the landscape "grain" relevant to a given organism may change over the course of its life stages [@addicott1987; @lima1996; @gardiner2010].
It also allows for the possibility that certain land cover types may be locally beneficial, but detremental at wider scales, indicating that the cover type may not constitute a completely usable habitat.

@@

For each landscape cover class, we used three scalar-on-function terms.
First, we used the proportion cover in the annuli surrounding each trap location (spatial effect of cover class).
Second, we used the average landcover in all annuli across days of the year (temporal effect of cover class). 
Finally, we used a tensor-product interaction of landscape cover and day of year (spatio-temporal interaction of cover class).
We used 10 basis dimensions for the spatial and temporal landscape smoothers, and 16 basis dimensions for the spatio-temporal interaction.
This was deemed adequate, as the effective degrees of freedom for all smoothing terms were far below the number of basis dimensions, indicating that no additional basis dimensions were needed (see Tables \@ref(tab:PteMelSmooth), \@ref(tab:ParDisSmooth), \@ref(tab:ParMoeSmooth), \@ref(tab:OpilioSmooth)).
To remove unimportant terms from the model, we used thin-plate regression splines with shrinkage, a continuous analog to stepwise model selection where weaker terms are completely removed from the model rather than reduced to a line [@marra2011].
The deviance residuals from each model were visually inspected for normality and equal variance [@hilbe2011; @wood2017].
Finally, we calculated the proportion of explained variance (R$^2$) for each set of terms [@nakagawa2013; @nakagawa2017].

@@

Functional linear regression is a special case of linear regression, where both the independent variable is predicted by a functional dependent variable, meaning that the slope is a function (vector) rather a fixed (scalar) value [@ramsay2004; @yen2014].
Therefore, the interpretation of a functional regression plot is very different from that of a simple linear regression plot.
Figure \@ref(fig:frExample)a shows a simple linear regression plot of a hypothetical relationship between trap _activity density_ (_y_) and a single predictor, _Proportion cover_ (_x_), which is measured in a circle of a fixed radius (_Distance_) surrounding each trap.
If we suppose that the relationship between _x_ and _y_ is similar at all radii at which _x_ is measured, then Figure \@ref(fig:frExample)b shows the resulting functional linear regression plot, where the slope ($\beta$) is the same at all distances.
In reality it is more likely that nearby cover will be more influential on activity density than far-away cover, which would result in a relationship similar to Figure \@ref(fig:frExample)c.
As functional regression takes any continuous predictor of _y_, it is also possible to use proportion cover (measured within a fixed radius) as a predictor of measurements taken across time, allowing the effect of proportion cover to vary over the course of the season (Figure \@ref(fig:frExample)d).
Finally, proportion cover over distance and time may be combined into a smoothed surface of effects, representing the spatio-temporal influence of landscape cover on activity density.

@@

Functional regression plots of landscape composition reveal which cover classes are acting as spillover sources or sinks of a given arthropod, and at what spatial scale.
Positive effects of nearby landscape cover classes indicate that arthropods are spilling over *from* it.
However, negative responses to nearby landscape cover classes can indicate one of two things:
1) It may indicate that arthropods are spilling over *into* it (i.e. net sink for spillover) or 2) it may indicate that no arthropods are available to spill over *from* it.
We consider 1) to be the more likely scenario, as it is unlikely that any one cover class is *completely* unoccupied.
<!-- As these data were collected from a single year, this model does not consider the next-generation fitness consequences of spillover. -->
For example, a cover class with low-quality habitat could act as an ecological trap [@galpern2017] if it acts as a net sink but never acts as a source of spill over.
Alternatively, a cover class could act as a sink by providing high-quality habitat at certain times of the year (e.g. egg laying/feeding), but act as a source at other times (e.g. during emergence).
For brevity in the following Results section, we refer to local- and landscape-level effects, where local indicates the effects of landscape composition at scales of <500m, whereas landscape indicates scales >500 m.
Similarly, we refer to early-, mid-, and late-season effects, where landscape composition affected abundance from June - July, July - August, or August - September, respectively.

@@

<!-- Percent signs are escaped in LaTeX, but this causes havoc if used in fig.cap of chunk options (OK if used in internal captions, I think). Yihui uses the following "text references" instead. -->
(ref:PteMelCap) Landscape influence on _Pterostichus melanarius_ activity-density. a) shows the effect of local trap location, and b-c) are functional regression plots of the proportion grassland and canola at distances away from each trap. d) is a functional regression plot over time, showing the effect of pulses at different times of the year. Lines and dots represent means, and bars and shaded regions represent 95 % confidence regions (1.96 x SE). Coloured regions represent early-, mid-, and late-season effects (blue = June, purple = July, red = August).

@@

Activity density of _P. distincta_ and _P. moesta_ was influenced by trap location and landscape composition (Tables \@ref(tab:ParDisLinear), \@ref(tab:ParDisSmooth), \@ref(tab:ParMoeLinear), \@ref(tab:ParMoeSmooth)), but landscape composition had relatively weak effects on both species ($R^2$: `r filter(r2Nakagawa2,grepl('Par',model),!grepl('s\\(',terms),terms!='trapLoc',terms!='sigmaD') %>% group_by(model) %>% summarize(s=sum(r2comp)) %>% pull(s) %>% paste(.,collapse=', ')`).
Unlike _P. melanarius_, both _Pardosa_ species had far lower activity-density in canola than any other cover type (Figures \@ref(fig:ParDisLandscape)a, \@ref(fig:ParMoeLandscape)a).
Activity-density of both _Pardosa_ species had a strong temporal and spatial component, although the temporal component was dominant for _P. distincta_, whereas the spatial component was dominant for _P. moesta_ (Figures \@ref(fig:ParDisSpatiotemporal) and \@ref(fig:ParMoeSpatiotemporal); both p`r with(pmoeSmTerms,p[Smoother=='s(E,N)'])`).

_P. distincta_ activity density was negatively affected by landscape-level pasture (Figure \@ref(fig:ParDisLandscape)b), while woodlands had a negative effect late in the season (Figure \@ref(fig:ParDisLandscape)c).
This suggests that large amounts of pasture are unsuitable habitat for _P. distincta_, and that they migrate into woodlands later in the season.
However, trap location and the temporal random effect (Figure \@ref(fig:ParDisSpatiotemporal)a) explained roughly 10 times more variance than landscape composition ($R^2$: `r round(sum(filter(r2Nakagawa2,model=='ParDisMod',terms=='trapLoc'|grepl('day',terms))$r2comp),3)` versus `r round(sum(filter(r2Nakagawa2,model=='ParDisMod',terms!='trapLoc',terms!='sigmaD',!grepl('s\\(',terms))$r2comp),3)`), meaning that while _P. distincta_ may be locally abundant in certain cover types, it is highly general in its overall habitat preferences.
<!-- The weak effect of canola crops (p>0.07) suggests that this species does not aggregate into these crops, meaning that its ecosystem service potential to farmers is limited. -->

>>>>

---
title: "Beneficial arthropod habitat useage changes within-season: implications for pest control services in croplands"
#Potential titles: 
# Non-crop features influence the abundance and timing of beneficial arthropod abundance
# 1) Cropland ecosystem services in space and time: Beneficial arthropod habitat use shifts within-season.
# 2) Beneficial arthropod habitat use shifts within-season: implications for pest control services in croplands
# 
author:
  - name: Samuel Robinson
    email: samuel.robinson@ucalgary.ca
    affiliation: University of Calgary
    footnote: 1
  - name: Diane Edwards
    email: diane@abienvironmental.ca
    affiliation: Ambrose University
  - name: Jess Vickruck
    email: jessica.vickruck@canada.ca
    affiliation: University of Calgary
  - name: Lincoln Best
    email: lincoln.best@oregonstate.edu
    affiliation: Oregon State University
  - name: Paul Galpern
    email: paul.galpern@ucalgary.ca
    affiliation: University of Calgary
address:
  - code: University of Calgary
    address: 2500 University Drive NW, Calgary, AB
  - code: Ambrose University
    address: 150 Ambrose Circle SW, Calgary, AB
  - code: Oregon State University
    address: 1500 SW Jefferson Way, Corvallis, OR
footnote:
  - code: 1
    text: "Corresponding Author"
date: "Spring 2020"
abstract: "Ground-dwelling arthropods are important generalist predators in agro-ecosystems, and can use non-crop features as overwintering habitats. However, it is unclear which types of landscape features constitute useful non-crop habitat, and at what spatial scale organisms gather resources at. Additionally, the same landscape feature may act as source of arthropods spillover at different times of the year, but this is rarely considered. We modeled the activity-density of four highly-abundant species of Canadian prairie arthropods, and used functional regression to simultaneously consider both their habitat preferences and the timing of sources and sinks. _Pterostichus melanarius_ (Coleoptera: Carabidae) and _Pardosa moesta_ (Arachnida: Lycosidae) appear to be attracted to canola (_Brassica napus_ L.) during the early summer, then disperse to grasslands, wetlands, and grassy road margins during the end of the summer. This suggests that grasslands and wetlands act as reservoir habitats for predatory arthropods, and that farmers should consider preserving existing habitat around croplands in order to maintain background levels of pest-control."

@@

# Introduction
<!-- DE: BIOCONTROL THEME NEEDS SET-UP IN INTRODUCTION -->

Arthropods are one of the most influential groups of animals in agricultural systems.
Pest arthropods can cause large reductions in crop yield [@oerke2005], which can result in negative downstream economic effects and food insecurity.
On the other hand, spillover of beneficial arthropods from adjacent non-crop patches can deliver increased crop yields by providing ecosystem services such as extra pollination or pest suppression [@losey2006; @klein2007; @garibaldi2013].
Seasonal "pulses" of food resources from crop fields act as a bottom-up driver of their associated arthropod communities, meaning that agricultural land can act as a source of nutrients and individuals for adjacent non-crop land [@diekotter2010; @galpern2017].
These important fluxes of mobile arthropods, both to crops and away from them, are key for managing both crop yield and biodiversity in agroecosystems.
Finding "win-win" practices is of particular importance to land managers and conservationists; that is, practices that result in better conservation outcomes while maintaining or improving crop yields.
<!-- Finally, agricultural land occupies 15 million km$^2$ (~36%) of land globally [@ramankutty2018], suggesting that there remains great potential to achieve conservation outcomes by adjusting agricultural practices. -->

Uncultivated semi-natural land (SNL) can act as important habitat for beneficial arthropods [@duelli2003], which may spill over into adjacent cropland.
Therefore, landscape management is a strategy to increase beneficial arthropod abundance in crops [@albrecht2010; @landis2000; @martin2020; but see @tscharntke2016; @karp2018].
For example, wild bee abundance in agricultural landscapes is largely driven by flower and nest availability [@roulston2011], and the effect of SNL on bee abundance and diversity is generally positive.
Predatory arthropods such as beetles or spiders can be limited by water, food, and egg-laying sites [@lovei1996; @purtauf2005; @gardiner2010], but the effect of SNL on their abundance and diversity is less consistent [@macfadyen2013; @shackelford2013] as some taxa may use SNL as a travel corridor or overwintering site, but are not completely dependent on it for reproduction.
Some taxa are not dependent on SNL at all, and thrive in cultivated landscapes, while others exclusively inhabit SNL and do not venture beyond the edge of cultivated fields.
For example, carabid beetles are thought to reproduce in field margins, while the adults feed in fields [@desender1988].
Habitat preference also depends on the life history of the organisms in question, such as the seasonal timing of emergence, feeding, and reproduction, as well as their traits such as dispersal capability.
Non-adjacent cropland can also influence arthropod abundance if the organisms are long-distance dispersers [@oberg2008].
<!-- Small patches of SNL can have a large effect if overall landscape lacks SNL (landscape contrast [@boetzl2020; @vanSchalkwyk2020]) -->
In this way, the spillover of beneficial arthropods into crops depends not only on the amount of adjacent SNL, but the specific habitat type, the time of the year, and the spatial scale at which an organism gathers resources.

@@

In this study, we used pitfall traps to determine how landscape composition affects the seasonal abundance of arthropods across a large Canadian prairie agroecosystem.
We considered the following hypotheses:
1. Untilled semi-natural land provides egg-laying and feeding areas for predatory arthropods, meaning that it should act as a source of arthropods during the early part of the season, and a sink during the later part of the season. 
2. Crops act as sources of food (pest insects) for predatory arthropods during the growing season. This should result in agricultural land becoming a sink for predators in the early part of the season as they migrate into the crop, and a source at the end of the season as they migrate out of the crop.
3. Crops may act as a temporary feeding site for predators. Therefore, crops may act as a local sink for arthropods, but will also be negatively associated at larger (landscape-level) scales.
Using a large pitfall trapping dataset, we related seasonal changes in arthropod abundance to landscape composition at multiple spatial scales.

@@

## Data collection
We used a set of `r sum(NtrapsPer$n)` pitfall traps installed in road margins (minimum of 5 m away from the road edge, `r with(NtrapsPer, sum(n[trapLoc=='ditch']))` traps) and in-field locations (`r with(NtrapsPer, sum(n[trapLoc!='ditch']))`) across southern Alberta, Canada in 2017 (Figure \@ref(fig:siteMap)).
The sites spanned a west-to-east gradient of four natural subregions, including foothills parkland, foothills fescue, mixedgrass, and dry mixedgrass [@ABRegions2006].
Traps were placed starting on `r format(as.Date(as.character(min(PteMelMod$tempTrap$startjulian)),format='%j'),format='%B %d')`, and collections ended on `r format(as.Date(as.character(max(PteMelMod$tempTrap$endjulian)),format='%j'),format='%B %d')`, with collection occurring continuously and traps being emptied every `r round(mean(PteMelMod$tempTrap$trapdays),1)` days on average (SD: `r round(sd(PteMelMod$tempTrap$trapdays),1)`). 
This resulted in `r nrow(PteMelMod$tempTrap)` unique collection events, taking place across a total of `r round(sum(PteMelMod$tempTrap$trapdays),1)` trapping days.
In-field traps were placed in canola crops (`r with(NtrapsPer, sum(n[trapLoc=='canola']))`), wetlands (`r with(NtrapsPer, sum(n[trapLoc=='wetland']))`), grassy field edges (`r with(NtrapsPer, sum(n[trapLoc=='pivot']))`), and remnant prairie grasslands (`r with(NtrapsPer, sum(n[trapLoc=='native']))`).
Traps in canola were installed at 25, 75, and 200 m along a transect heading away from the nearest non-crop feature (wetland, grassy field edges, or remnant prairie), while the trap at 0 m was installed in the non-crop feature itself.
We used 582 mL Solo® cups buried up to the rim and partially filled with propylene glycol, with 2 cm wire mesh mounted over the rim to prevent vertebrates from falling into the traps.
Specimens were identified to species using appropriate taxonomic literature [@dondale1990; @edgar1990; @vogel2004; @yigit2007].

We used counts of organisms in our pitfall traps as proxies of ecosystem service provision.
However, increased counts of organisms in pitfall traps can represent higher activity levels (same number of organisms but more mobile) or higher abundances in the vicinity of the trap.
This makes it impossible to disentangle arthropod activity from density using single traps [@lang2000; @zhao2013], meaning that counts in the traps represent the "activity-density" of a given organism.
Despite this drawback, activity-density is generally positively related to ecosystem services, at least in carabids [@trichard2014; @boetzl2018; @gonzalez2020], making it acceptable for our study.

```{r siteMap, echo=FALSE, fig.cap='Location of pitfall traps during 2017, and classified land cover. Inset map shows study location within Alberta. Roadside traps were installed in road margins, while in-field traps were installed directly in canola fields or in adjacent wetland or grassy field edges.', out.width='100%'}  
  include_graphics('../figures/maps/PF2017Map.png',dpi=NA)
```

We used four highly-abundant species of predatory arthropods found in the pitfall traps, as common species are often more important for ecosystem service provision than rare species [@winfree2015].
_Pterostichus melanarius_ is an introduced predatory ground beetle (Coleoptera: Carabidae) that is a wide-ranging generalist known to prey upon crop pests [@carcamo1994; @larsen2003; @busch_Msc].
It is commonly found in untilled grasslands [@purtauf2005] and may use grassy field edges as larval habitat [@desender1988].
_Pardosa_ are a genus of wolf spiders (Araneae: Lycosidae) that use an active wandering predation strategy [@young1990], and are common across the Canadian prairies.
Both _Pardosa distincta_ and _Pardosa moesta_ are found in a wide variety of habitats, but are common in ungrazed pastures [@dondale1990; @carcamo2014].
_P. distincta_ is also commonly found in disturbed environments [@collins1996; @wade2010], while _P. moesta_ appear to prefer less frequent disturbances [@graham2003; @kowal2011; but see @dondale1990].
Other _Pardosa_ are known to use road margins and wooded areas as habitat [@buddle2000; @drapela2011].
Finally, _Phalangium opilio_ is an introduced harvestman (Opiliones: Phalangiidae) that is found in drier disturbed areas [@bragg2009].
Juveniles hatch from eggs during the spring, becoming adults during the summer, and eggs are laid during the fall [adults do not overwinter; @bragg2009].
They are commonly found in human-altered landscapes [@muster2014; @vandepoel2015Msc] and are nocturnal generalist hunters and scavengers [@halaj2000; @allard2005; @allard2005b], mainly eating small soft-bodied invertebrates [@acosta2007].

@@

Functional regression [@ramsay2004; @yen2014] was used to incorporate landscape information at different distances, allowing assessment of both local and regional landscape composition [@galpern2020].
Scalar-on-function regression is a special type of linear regression model ($y=X\beta + e$), where the columns of the model matrix $X$ contain some continuous predictor of the scalar $y$, and the values of the coefficients ($\beta$) are modeled as a smooth function $f(x)$ of the predictors.
<!-- As an example of scalar-on-function regression, biomass in a plot (scalar) could be modeled as a smoothed function of monthly average temperatures at that plot (function), meaning that the effect of monthly temperatures would also be a function of time. -->
In our case, the proportion of landscape cover ($X$) within each annulus surrounding the trap is a function of distance away from the trap, meaning that the coefficients are a smoothed function of distance ($\beta_i = f(distance_i)$), and represent the additive effect of a given type of landscape cover at distances away from the trap (i.e. $X\beta$).
This allows for the possibility that the size of the landscape "grain" relevant to a given organism may change over the course of its life stages [@addicott1987; @lima1996; @gardiner2010].
It also allows for the possibility that certain land cover types may be locally beneficial, but detremental at large scales, indicating that the cover type may not constitute a completely usable habitat.

@@

For each landscape cover class, we used three scalar-on-function terms.
First, we used the proportion cover in the annuli surrounding each trap location (spatial effect of cover class).
Second, we used the average proportion cover across days of the year (temporal effect of cover class). 
Finally, we used a tensor-product interaction of landscape cover and day of year (spatio-temporal interaction of cover class).
We used 10 basis dimensions for the spatial and temporal landscape smoothers, and 16 basis dimensions for the spatio-temporal interaction.
The effective degrees of freedom for all smoothing terms were far below the number of basis dimensions, indicating that no additional basis dimensions were needed (see Tables \@ref(tab:PteMelSmooth), \@ref(tab:ParDisSmooth), \@ref(tab:ParMoeSmooth), \@ref(tab:OpilioSmooth)).
To remove unimportant terms from the model, we used thin-plate regression splines with shrinkage, a continuous analog to stepwise model selection where weaker terms are completely removed from the model rather than reduced to a line [@marra2011].
The deviance residuals from each model were visually inspected for normality and equal variance [@hilbe2011; @wood2017].
Finally, we calculated the proportion of explained variance (R$^2$) for each set of terms [@nakagawa2013; @nakagawa2017].

@@

Functional linear regression is a special case of linear regression, where both the independent variable is predicted by a functional dependent variable, meaning that the slope is a function (vector) rather a fixed (scalar) value [@ramsay2004; @yen2014].
Therefore, the interpretation of a functional regression plot is very different from that of a simple linear regression plot.
Figure \@ref(fig:frExample)a shows a simple linear regression plot of a hypothetical relationship between trap _activity-density_ (_y_) and a single predictor, _Proportion cover_ (_x_), which is measured in a circle of a fixed radius (_Distance_) surrounding each trap.
If we suppose that the relationship between _x_ and _y_ is similar at all radii at which _x_ is measured, then Figure \@ref(fig:frExample)b shows the resulting functional linear regression plot, where the slope ($\beta$) is the same at all distances.
In reality it is more likely that nearby cover will be more influential on activity-density than far-away cover, which would result in a relationship similar to Figure \@ref(fig:frExample)c.
As functional regression takes any continuous predictor of _y_, it is also possible to use proportion cover (measured within a fixed radius) as a predictor of measurements taken across time, allowing the effect of proportion cover to vary over the course of the season (Figure \@ref(fig:frExample)d).
Finally, proportion cover over distance and time may be combined into a smoothed surface of effects, representing the spatio-temporal influence of landscape cover on activity-density.

@@

Functional regression plots of landscape composition reveal which cover classes are acting as spillover sources or sinks of a given arthropod, and at what spatial scale.
Positive effects of nearby landscape cover classes indicate that arthropods are spilling over *from* it.
However, negative responses to nearby landscape cover classes can indicate one of two things:
1) It may indicate that arthropods are spilling over *into* it (i.e. net sink for spillover) or 2) it may indicate that no arthropods are available to spill over *from* it.
We consider 1) to be the more likely scenario, as it is unlikely that any one cover class is *completely* unoccupied.
<!-- As these data were collected from a single year, this model does not consider the next-generation fitness consequences of spillover. -->
For example, a cover class with low-quality habitat could act as an ecological trap [@galpern2017] if it acts as a net sink but never acts as a source of spill over.
Alternatively, a cover class could act as a sink by providing high-quality habitat at certain times of the year (e.g. egg laying/feeding), but act as a source at other times (e.g. during emergence).
In the Results section, we refer to local- and landscape-level effects, where local indicates the effects of landscape composition at scales of <500 m, whereas landscape indicates scales >500 m.
Similarly, we refer to early- and late-season effects, where landscape composition affected activity-density on June 20th or August 20th, respectively.

@@

<!-- Percent signs are escaped in LaTeX, but this causes havoc if used in fig.cap of chunk options (OK if used in internal captions, I think). Yihui uses the following "text references" instead. -->
(ref:PteMelCap) Landscape influence on _Pterostichus melanarius_ activity-density. a) shows the effect of local trap location, and b-c) are functional regression plots of the proportion grassland and canola at distances away from each trap. d) is a functional regression plot over time, showing the effect of pulses at different times of the year. Lines and dots represent means, and bars and shaded regions represent 95 % confidence regions (1.96 x SE). Coloured regions represent early- and late-season effects (blue = June, red = August).

@@

Activity-density of _P. distincta_ and _P. moesta_ was influenced by trap location and landscape composition (Tables \@ref(tab:ParDisLinear), \@ref(tab:ParDisSmooth), \@ref(tab:ParMoeLinear), \@ref(tab:ParMoeSmooth)), but landscape composition had relatively weak effects on both species ($R^2$: `r filter(r2Nakagawa2,grepl('Par',model),!grepl('s\\(',terms),terms!='trapLoc',terms!='sigmaD') %>% group_by(model) %>% summarize(s=sum(r2comp)) %>% pull(s) %>% paste(.,collapse=', ')`).
Unlike _P. melanarius_, both _Pardosa_ species had far lower activity-density in canola than any other cover type (Figures \@ref(fig:ParDisLandscape)a, \@ref(fig:ParMoeLandscape)a).
Activity-density of both _Pardosa_ species had a strong temporal and spatial component, although the temporal component was dominant for _P. distincta_, whereas the spatial component was dominant for _P. moesta_ (Figures \@ref(fig:ParDisSpatiotemporal) and \@ref(fig:ParMoeSpatiotemporal); both p`r with(pmoeSmTerms,p[Smoother=='s(E,N)'])`).

_P. distincta_ activity-density was negatively affected by landscape-level pasture (Figure \@ref(fig:ParDisLandscape)b), while woodlands had a negative effect late in the season (Figure \@ref(fig:ParDisLandscape)c).
This suggests that large amounts of pasture are unsuitable habitat for _P. distincta_, and that they migrate into woodlands later in the season.
However, trap location and the temporal random effect (Figure \@ref(fig:ParDisSpatiotemporal)a) explained roughly 10 times more variance than landscape composition ($R^2$: `r round(sum(filter(r2Nakagawa2,model=='ParDisMod',terms=='trapLoc'|grepl('day',terms))$r2comp),3)` versus `r round(sum(filter(r2Nakagawa2,model=='ParDisMod',terms!='trapLoc',terms!='sigmaD',!grepl('s\\(',terms))$r2comp),3)`), meaning that while _P. distincta_ may be locally abundant in certain cover types, it is highly general in its overall habitat preferences.
<!-- The weak effect of canola crops (p>0.07) suggests that this species does not aggregate into these crops, meaning that its ecosystem service potential to farmers is limited. -->
